Validation of quantifying infant all-day restrained experiences at home using inertial sensors

### 2025-04-21 manuscript_restrained_classification validation.qmd
modified introduction

However, restraint does not always reduce infant learning opportunities. Sometimes caregiver carrying infants or putting infants in devices can rescue infants from their body disadvantages, particularly during earlier stages of development. For example, young infants who cannot sit independently struggle to interactive with objects because they cannot visually examine and manually manipulate the object at the same time [@Soska2010q]. However, being seated in supportive devices helps stablize infant posture, free infant hands and allow for multimodal object interaction (XXX). Similarly, infant body height limits what they can see on their own. The situation is even worse for infants who cannot stand up because they see much fewer faces when they are in prone position [@Franchak2018i]. However, being held in a forward-facing carrier or seated in highchairs can elevate infants' viewpoint and increase opportunities for face-to-face interaction [@Yu2013h; @Kretch2015p; @Franchak2024d]. These examples demonstrate that restraint may provide infants with uncommon learning materials that they would not otherwise obtained due to their developmental constraint. These examples further imply that the effects of restraint may vary depending on infants’ age and motor abilities, reinforcing the need to precisely quantify restraint across developmental stages and examine how it changes with development.

@@

Does infant restrained time change with age? Only a handful of studeis have answered this question, but their findings are mixed as well. @Airaksinen2024z, using inertial measurement sensors (IMU) and machine learning algorithms, found that infants from Finland experienced *increased* holding and carrying time from 4 to 18 months of age. On the contrary, @Carson2022q, using parental retrospective report and longitudinal design, found that infants from Canada experienced *decreased* overall restrained time from 2 to 6 months of age. @Franchak2024v, using EMA and longitudinal design, found that infants from the US experienced *decreased* time held by caregivers from 10 to 13 months, yet the time restrained by furniture or devices did not change with age. The discrepancies in findings may be attributed to differences in the targeted behaviors (i.e., whether to include device restraint or not, and which devices to include), study design (i.e., cross-sectional vs. longitudinal), cultural contexts, as well as research methods. To address the current discrepancies in findings, the first step is to develop an accurate and reliable system that examines infant general restraint experiences, which can be further applied to different age ranges and culture contexts. Among the available methods, we propose IMU is the most objective, unobtrusive, and efficient way to caputure whole-day experiences in natural settings. But previous studies that have used IMU on infants only looked at holding behavior, excluding other extensive restrained practices such as device use [@Airaksinen2024z; @Franchak2024i]. Thus, a new customized machine learning algorithm is needed to identify a broader range of restraint experiences.

@@

Given the limitations in current methods and the potential of IMU technology, the overall goal of this study is to evaluate the validity of using IMUs to precisely quantify infant full-day restrained experiences at home. Despite previous validation of IMU system on identifying infant position and caregiver holding, identifying all restraint practices on infants is still challenging. Consider an infant sitting on a couch independently versus an infant sitting on a highchair restrained, or an infant lying supine unrestrained versus an infant being tucked in cradle restrained. The infants are in same body position and are very likely to show similar movement patterns, but the context is different, differentiating the restrained status. Furthermore, device restraint might be more difficult to detect than caregiver restraint because carrying and holding tend to generate dramatic movement (e.g., the infant moving in an adult speed). Thus, in the current study, we develop and validate the use of new machine-learning models that identify both caregiver restraint and device restraint, classifying infant moment-to-moment status as being restrained or unrestrained.

The models are tested by a total of 143 sessions from 66 infants who were aged from 7 to 14 months, covering a wide range from pre-sitting infants to walking infants. We sent caregivers special leggings with four IMUs embedded and a camera to record synchronized infant activities. Across the sessions, IMUs recorded an average of XX hours (*SE* = XX) of movement data, and the camera recorded the first 1.5 hours into the session. From the videos, human coders identify infant restrained status as ground truth for model training and testing. The **first aim** of the current study is to train the supervised machine-learning models of infant restrained status and evaluate the models’ accuracy using leave-one-out cross validation method (XXX) for the first 1.5 hours.

>>>>

However, restraint does not always reduce infant learning opportunities. Sometimes caregiver carrying infants or putting infants in devices can rescue infants from their body disadvantages, particularly during earlier stages of development. For example, young infants who cannot sit independently struggle to interactive with objects because they cannot visually examine and manually manipulate the object at the same time [@Soska2010q]. However, being seated in supportive devices helps stabilize infant posture, free infant hands and allow for multimodal object interaction (XXX). Similarly, infant body height limits what they can see on their own. The situation is even worse for infants who cannot stand up because they see much fewer faces when they are in prone position [@Franchak2018i]. However, being held in a forward-facing carrier or seated in highchairs can elevate infants' viewpoint and increase opportunities for face-to-face interaction [@Yu2013h; @Kretch2015p; @Franchak2024d]. These examples demonstrate that restraint may provide infants with uncommon learning materials that they would not otherwise obtained due to their developmental constraint. These examples further imply that the effects of restraint may vary depending on infants’ age and motor abilities, reinforcing the need to precisely quantify restraint across developmental stages and examine how it changes with development.

@@

Does infant restrained time change with age? Only a handful of studies have answered this question, but their findings are mixed as well. @Airaksinen2024z, using inertial measurement sensors (IMU) and machine learning algorithms, found that infants from Finland experienced *increased* holding and carrying time from 4 to 18 months of age. On the contrary, @Carson2022q, using parental retrospective report, found that infants from Canada experienced *decreased* overall restrained time from 2 to 6 months of age. @Franchak2019v, using EMA, found that infants from the US experienced *decreased* time held by caregivers from 3 to 12 months. The author further found 10- to 13-month-olds also experienced *decreased* held time, and *increased* overall unrestrained time [@Franchak2024v]. The discrepancies in findings may be attributed to differences in the targeted behaviors (i.e., whether to include device restraint or not, and which devices to include), study design (i.e., cross-sectional vs. longitudinal), cultural contexts, as well as research methods. To address the current discrepancies in findings, the first step is to develop an accurate and reliable system that examines infant general restraint experiences, which can be further applied to different age ranges and culture contexts. Among the available methods, we propose IMU is the most objective, unobtrusive, and efficient way to capture whole-day experiences in natural settings. But previous studies that have used IMU on infants only looked at holding behavior, excluding other extensive restrained practices such as device use [@Airaksinen2024z; @Franchak2024i]. Thus, a new customized machine learning algorithm is needed to identify a broader range of restraint experiences.

@@

Given the limitations in current methods and the potential of IMU technology, the overall goal of this study is to evaluate the validity of using IMU to precisely quantify infant full-day restrained experiences at home. Despite previous validation of IMU system on identifying infant position and caregiver holding, identifying all restraint practices on infants is still challenging. Consider an infant sitting on a couch independently versus an infant sitting on a highchair restrained, or an infant lying supine unrestrained versus an infant being tucked in cradle restrained. The infants are in same body position and are very likely to show similar movement patterns, but the context is different, differentiating the restrained status. Furthermore, device restraint might be more difficult to detect than caregiver restraint because carrying and holding tend to generate dramatic movement (e.g., the infant moving in an adult speed). Thus, in the current study, we develop and validate the use of new machine-learning models that identify both caregiver restraint and device restraint, classifying infant moment-to-moment status as being restrained or unrestrained.

The models are tested by a total of 143 sessions from 66 infants who were aged from 7 to 14 months, covering a wide range from pre-sitting infants to walking infants. We sent caregivers special leggings with four IMU embedded and a camera to record synchronized infant activities. Across the sessions, IMU recorded an average of XX hours (*SE* = XX) of movement data, and the camera recorded the first 1.5 hours into the session. From the videos, human coders identify infant restrained status as ground truth for model training and testing. The **first aim** of the current study is to train the supervised machine-learning models of infant restrained status and evaluate the models’ accuracy using leave-one-out cross validation method (XXX) for the first 1.5 hours.
